\section{Conclusion and Future Work}
In this paper, we proposed a generalized explanation engine called \sys that serves as the basis for 
future interactive visualizations systems that can explore, visualization, and explain.
\sys poses explanation generation as a planning problem over a language of data transformations in order to consider datasets with combinations of systematic errors.
It also expresses existing error-specific approaches---constraint-based, statistical, and demonstration-based data cleaning---within a single, generic search framework. 
The reason for this generic approach is to provide the flexibility to easily specifying different types of output anomalies, and mix-and-match the types of errors to consider.
Although this search problem is theoretically very challenging, our results suggest that borrowing from recent advances in planning and optimization is a fruitful direction.  

This work primarily focused on the core explanation engine, and sketched ways that users may express anomalies througha visualization interface.
Our immediate plans are to develop an interaction language to express a comprehensive class of anomalies that can be found in BI and machine learning model visualizations. 
This will also require work to characterize failure modes and provide high-level tools to debug such cases.  
We are also hopeful that the compact codebase (<200 LOC for the core search and learning algorithms) can enable more rapid development of specialized data cleaning systems for novel domains and error conditions.  




% The prevailing wisdom in the design of data cleaning algorithms is to exploit the details of specific problem rather than considering the most general cases, and our experiments suggest that this a general framework like \sys can achieve parity in terms of accuracy.
% While the serial implementation of \sys can be much slower than the competitor specialized frameworks, \sys can be efficiently distributed to significantly reduce the gap.
% 
% These results should be considered a proof-of-concept that such a data cleaning \sys can be built around the recent results in AI. 
% However, to us, these results are still counter-intuitive, and raise a number of speculative questions for the future: (1) are specialized systems overly engineered for worst-case guarantees and perhaps real-datasets are not that pathalogical, (2) maybe the benchmarks that we consider in data cleaning are too easy to brute-force, (3) what are the failure modes and corner cases of \sys in real data.
% We hope to consider these problems in future work, as well as extending the system to novel settings.
% In particular, we are interested in \sys as a middleware layer for data visualization.
% A user can manipulate data in a visual UI and these manipulations can be translated into a quality function.
% 
